---
title: "Midpoint Deliverable Report"
output: html_document
---
```{r, echo = FALSE}

health_df <- read.csv("scripts/data/us-deaths.csv", stringsAsFactors = FALSE)
source("scripts/summary.R")
source("scripts/tableFunction.R")
source("scripts/bargraph.R")
source("scripts/scatterplot.R")
source("piechart.R")
```

library(“dplyr”)
library(“ggplot2”)


## Introduction
With our current data, we hope to observe trends among Covid-19, Pneumonia, and Influenza and see how they affect death rates in each state. 

The main purpose of using this data is to see where and who the deadly covid-19 pandemic is 

In our project, we are exploring the death rates involving Covid-19, Influenza and Pneumonia across the US.

The data analyzed in this report is from [CDC](https://data.cdc.gov/NCHS/Provisional-COVID-19-Death-Counts-by-Sex-Age-and-S/9bhg-hcku?fbclid=IwAR3VbsTWlBjY4w8Rz8rpnO3Bv1w0zDrHkhM2hB5OXDrfvLFudwF1f1Kkbn8).

## Summary Information
``` {r echo = FALSE, include = FALSE}
library(“knitr”)
kable(get_summary_info(health_df))

```
From 2/1/20 to 5/16/20, the state with the most deaths due to COVID 19 was found to be
`r deadliest_covid_state`. The state with the most overall deaths was found to be
`r deadliest_state`. There are `r wa_covid_deaths` deaths in Washington caused by COVID 19.
The age group that has the most deaths in the US due to COVID 19 was found to be those who were `r age_most_affected`. Finally, the COVID 19 death percentage in the US is calculated to be `r covid_death_percentage`.

## Aggregated Table

```{r kable, echo = FALSE}
library(knitr)
health_table <- get_table(health_df)
knitr::kable(heath_table, 
            col.names = c("State", "% of Covid-19 deaths", "% of Pneumonia deaths", "% of Influenza deaths"))
```
This table displays the top 10 states with the percentages of deaths that occurred from February 1st 2020 to May 16 2020 and the US total's percentages as well. These percentages were derived from dividing the cause of death numbers with the total deaths. (It reveals what percentage of the deaths were caused by what illness.) These percentages are grouped by location/state and include people from all gender and ages. The table is ordered in descending order of Covid-19 death percentages. In this way, it is easier for us to observe which locations had the most Covid death ratios and compare it to the other causes of death.
The table reveals to us that the highest percentages of Covid deaths were in New York City. About 40% of the deaths that occurred in the timeframe were due to Covid. Likewise, in most states, Covid is the leading cause of death among the three categories. However, for the US total, Pneumonia is the leading cause of death. Also this table illustrates that the highest Covid death percentages are more towards the Eastern side of America.
