---
title: "index"
output: html_document
---
```{r, echo = FALSE}
source("scripts/tableFunction.R")

```

## Introduction

## Summary Information

## Aggregated Table

```{r kable, echo = FALSE}
library("knitr")
knitr::kable(heath_table, 
            col.names = c("State", "% of Covid-19 deaths", "% of Pneumonia deaths", "% of Influenza deaths"))
```

This table displays the percentages of deaths that occurred from February 1st 2020 to May 16 2020. These percentages were derived from dividing the cause of death numbers with the total deaths. (It reveals what percentage of the deaths were caused by what illness.) These percentages are grouped by location/state and include people from all gender and ages. The table is ordered in descending order of Covid-19 death percentages. In this way, it is easier for us to observe which locations had the most Covid death ratios and compare it to the other causes of death.
The table reveals to us that the highest percentages of Covid deaths were in New York City. About 40% of the deaths that occurred in the timeframe were due to Covid. Likewise, in most states, Covid is the leading cause of death among the three categories. 

## Chart 1

## Chart 2

## Chart 3